---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# RabbitR

<!-- badges: start -->
<!-- badges: end -->

This is a teaching-oriented package for univariate linear mixed models solved using Bayesian theory. The package includes visualization and characterization of marginal posterior distributions of the estimates.

## Installation

You can install the development version of RabbitR from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("marinamartinezalvaro/RabbitR")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(RabbitR)

```

In this example, we are going to use DataIMF from RabbitR package.The DataIMF dataset (502 observations across 8 variables) comes from three generations of a divergent selection experiment for intramuscular fat in rabbits.The data are based on perirenal fat (PFat), intramuscular fat (IMF) and loin pH of rabbits belonging to both sexes (Sex), with an specific live weight (LW). The data have been taken at different seasons (AE, 2 levels) and rabbits were born at different parity orders (OP),and have a random litter effect (c).This is how DataIMF looks like:

```{r DataIMF}
data(DataIMF)
head(DataIMF)

```


We aim to explore the impact of Sex on IMF and PFat traits. Additionally, we recognize the potential influence of the season when data were collected, the parity order of the animals, their LW, and their common litter effects on IMF, which necessitates adjustment. In this context, Sex is considered our treatment of interest. Other fixed effects that require correction but are not our primary focus are categorized as Noise effects (AE, OP). Live Weight (LW) is a continuous trait and will thus be treated as a covariate. Effects from common litter will be modeled as random effects.

The RabbitR package supports univariate analysis, applying the same univariate model across multiple traits:

IMF = m + Sex + AE + OP + AE*OP + b·LW + Rand(c) + e

PFat = m + Sex + AE + OP + AE*OP + b·LW + Rand(c) + e

To carry out our analysis, we'll utilize the CreateParam, Bunny, and Bayes functions. This workflow covers creating a parameter file (CreateParam), executing general linear models and producing posterior distributions (Bunny), and performing inferences (Bayes). For beginners or those less familiar with R, interactive versions are available: iCreateParam for CreateParam and iBayes for Bayes. Additionally, the Rabbit function merges iCreateParam, Bunny, and iBayes into a comprehensive, interactive tool, ideal for educational purposes.

Lets start creating our parameter file. 
To initiate our analysis, we'll create a parameter file using a CreateParam function. User can specify Traits, Treatments, Noise, covariates, or random effects. You can define these either by their column names in the data file (hTrait, hTreatment, hNoise, hCov, hRand) or by their positions (pTrait, pTreatment, pNoise, pCov, pRand), useful for analyzing multiple traits. The askCompare argument lets you choose how to contrast different treatment levels (e.g., sex1 and Sex2) - as a difference (askCompare="D") or a ratio (askCompare="R").

```{r CreateParam, echo = TRUE}

param_list<-CreateParam(
  file.name = "DataIMF",
  na.codes=c("99999"),
  hTrait = c("IMF", "PFat"),
  hTreatment = "Sex",
  askCompare="D",
  hNoise = c("AE", "OP"),
  hInter=matrix(c("AE","OP"), nrow=1),
  ShowInter=c("T"),
  hCov = c("LW"),
  hRand = "c")

```

Once the parameter file is ready, we can proceed to fit our model with Bunny

```{r Bunny, echo = TRUE}

bunny_results <- Bunny(params = param_list, Chain = FALSE)

```


Function Bunny creates a list containing the samples of the posterior chains for the model mean, means of treatment levels, contrats between those levels, covariates, variance components (residual and random). The bunny output will serve as input for function Bayes, together with my param_list object.

Although posterior chains might seem complex at first, they offer us the ability to make multiple inferences from the entire distribution, not just single-point estimates-making our analysis more engaging. The function simplifies in-depth analysis of posterior distributions from Bayesian mixed models. For instance, we calculate the probability of the difference between Sex1 and Sex2 exceeding critical values (0.05 for IMF and 1 for PFat) and obtain guaranteed values for treatment means, covariates, and contrasts at a 90% confidence level. Enabling plot=TRUE visually presents these distributions, with dashed lines at 0 and R clarifying the positions of P0, PR, and PS. 

```{r Bayes, echo = TRUE}

inferences <- Bayes(
   params = param_list,
   bunny = bunny_results,
   HPD = 0.95,
   K = TRUE, # Compute a guaranteed value of the estimate
   probK = 0.90, # With 90% probability
   PR = TRUE, # Compute the probability of the posterior chain being greater than a relevant value
   R = c(0.05, 1), # Assuming two traits with relevant values specified
   PS = TRUE, # Compute probability of similarity
   SaveTable = TRUE, # Save detailed inferences in a CSV file
   plot = TRUE) # Additionally, generate plots for contrasts)

```

This package makes extensive use  of `MCMCglmm` R package. We acknowledge the work by Jarrod D. Hadfield in this area, as detailed in the following reference:

Hadfield, J. D. (2010). MCMC methods for multi-response generalized linear mixed models: the MCMCglmm R package. *Journal of Statistical Software*, 33, 1-22. Available at [Journal of Statistical Software](https://www.jstatsoft.org/article/view/v033i02).
